---
title: " TUTORIAL EXERCISES @ http://r4ds.had.co.nz/data-visualisation.html"
author: "Manojit Sarkar"
date: "7 July 2018"
output: html_document
---

##Post exercise library calls
```{r}

library(tidyverse)
```


##3.2.4 Exercises

1)Run ggplot(data = mpg). What do you see?

```{r}
ggplot(mpg)
```

Ans: No graph plot shown.

2)How many rows are in mpg? How many columns?

```{r dimension mpg}
dim(mpg)
```
Ans : 234 rows and 11 coloumns.

3)What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg
```

Ans :f = front-wheel drive, r = rear wheel drive, 4 = 4wd (what drive the car is )

4)Make a scatterplot of hwy vs cyl.

1st code :
```{r}
ggplot(mpg,aes(x=hwy,y=cyl))+geom_point()
```

2nd Code:
```{r}
ggplot(mpg)+geom_point(mapping = aes(x=hwy,y=cyl))
```



5)What happens if you make a scatterplot of class vs drv? Why is the plot not useful?


```{r}
ggplot(mpg,aes(x=class,y=drv))+geom_point()
```

Ans: The plot is not useful as the variables used are categorical variables.


## 3.3.1 Exercises
1)What’s gone wrong with this code? Why are the points not blue?


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```


Ans:Since colour = "blue" was included within the mapping argument, it was treated as an aesthetic (a mapping between a variable and a value). The expression, color="blue", treats "blue" as a variable with only one value: "blue". If this is confusing, consider how colour = 1:234 or colour = 1 would be interpreted by aes(). 
@ ans from  https://github.com/jrnold/r4ds-exercise-solutions/blob/master/visualize.Rmd

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```


2)Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

Ans there are many ways to know this 

1st way 

```{r}
sapply(mpg,class)
```

2nd way 

```{r}
? mpg 
## and then check the documentation
```

<UL>
<h1> Categorical Variables</h1>
<li>manfacturer</li>
<li>model</li>
<li>trans</li>
<li>drv</li>
<li>fl</li>
<li>class</li>

</UL>

<UL>
<h1> Continous Variables</h1>
<li>displ</li>
<li>hwy</li>
<li>year<li>
<li>cty</li>
</UL>

3)Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

## COLOR
```{r}
ggplot(mpg,aes(x=cty,y=displ,color=hwy))+geom_point()
```

## SIZE
```{r}
ggplot(mpg,aes(x=cty,y=displ,size=hwy))+geom_point()
```

## SHAPE
```{r}
#ggplot(mpg,aes(x=cty,y=displ,shape=year))+geom_point()
```

 Continous variable cannot be mapped to shape
 
## Color with categorical variable
 
```{r}
ggplot(mpg,aes(x=cty,y=displ,color=class))+geom_point()
```
Ans :For continous variable there was same colour with different intensity , while for categorical variables there are different colours for each type.


## Size with categorical variable

```{r}
ggplot(mpg,aes(x=cty,y=displ,size=fl))+geom_point()
```

A warning with size is not advised for discrete variables.


```{r}
ggplot(mpg,aes(x=cty,y=displ,shape=fl))+geom_point()
```
 It display the grap as against continous variable. But shape palette can display a maximum of 6 shapes.

3)What happens if you map the same variable to multiple aesthetics?




What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?


