---
title: ' TUTORIAL EXERCISES @ http://r4ds.had.co.nz/data-visualisation.html'
author: "Manojit Sarkar"
date: "7 July 2018"
output:
  html_document: default
  pdf_document: default
---

##Pre exercise library calls
```{r}

library(tidyverse)
```


##3.2.4 Exercises

1)Run ggplot(data = mpg). What do you see?

```{r}
ggplot(mpg)
```

Ans: No graph plot shown.

2)How many rows are in mpg? How many columns?

```{r dimension mpg}
dim(mpg)
```
Ans : 234 rows and 11 coloumns.

3)What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg
```

Ans :f = front-wheel drive, r = rear wheel drive, 4 = 4wd (what drive the car is )

4)Make a scatterplot of hwy vs cyl.

1st code :
```{r}
ggplot(mpg,aes(x=hwy,y=cyl))+geom_point()
```

2nd Code:
```{r}
ggplot(mpg)+geom_point(mapping = aes(x=hwy,y=cyl))
```



5)What happens if you make a scatterplot of class vs drv? Why is the plot not useful?


```{r}
ggplot(mpg,aes(x=class,y=drv))+geom_point()
```

Ans: The plot is not useful as the variables used are categorical variables.


## 3.3.1 Exercises
1)What’s gone wrong with this code? Why are the points not blue?


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```


Ans:Since colour = "blue" was included within the mapping argument, it was treated as an aesthetic (a mapping between a variable and a value). The expression, color="blue", treats "blue" as a variable with only one value: "blue". If this is confusing, consider how colour = 1:234 or colour = 1 would be interpreted by aes(). 
@ ans from  https://github.com/jrnold/r4ds-exercise-solutions/blob/master/visualize.Rmd

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```


2)Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

Ans there are many ways to know this 

1st way 

```{r}
sapply(mpg,class)
```

2nd way 

```{r}
? mpg 
## and then check the documentation
```

<UL>
<h1> Categorical Variables</h1>
<li>manfacturer</li>
<li>model</li>
<li>trans</li>
<li>drv</li>
<li>fl</li>
<li>class</li>

</UL>

<UL>
<h1> Continous Variables</h1>
<li>displ</li>
<li>hwy</li>
<li>year<li>
<li>cty</li>
</UL>

3)Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

## COLOR
```{r}
ggplot(mpg,aes(x=cty,y=displ,color=hwy))+geom_point()
```

## SIZE
```{r}
ggplot(mpg,aes(x=cty,y=displ,size=hwy))+geom_point()
```

## SHAPE
```{r}
#ggplot(mpg,aes(x=cty,y=displ,shape=year))+geom_point()
```

 Continous variable cannot be mapped to shape
 
## Color with categorical variable
 
```{r}
ggplot(mpg,aes(x=cty,y=displ,color=class))+geom_point()
```
Ans :For continous variable there was same colour with different intensity , while for categorical variables there are different colours for each type.


## Size with categorical variable

```{r}
ggplot(mpg,aes(x=cty,y=displ,size=fl))+geom_point()
```

A warning with size is not advised for discrete variables.


```{r}
ggplot(mpg,aes(x=cty,y=displ,shape=fl))+geom_point()
```
 It display the grap as against continous variable. But shape palette can display a maximum of 6 shapes.

3)What happens if you map the same variable to multiple aesthetics?

multi aesthetics on continous variable

```{r}
ggplot(mpg,aes(x=cty,y=displ,color=hwy,size=hwy))+geom_point()
```

yes it can be represented with color and size for continous variables.

multi aesthetics on caterogical variables.

```{r}
ggplot(mpg,aes(x=cty,y=displ,color=class,size=class))+geom_point()
```
Ans With same categorial variables can be plotted but with warnings.




4)What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}
?geom_point

ggplot(mpg,aes(x=cty,y=displ,shape=fl))+geom_point(stroke=4)
```

stroke alters the sizes of shapes


5)What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?

```{r}
ggplot(mpg,aes(x=displ,y=hwy,color=displ<5))+geom_point()
```
   
   It changes the values into boolean i.e true and false ; true / false depicting different colours each and is ploted to the graph.
   
   
   
##3.5.1 Exercises
   
1)What happens if you facet on a continuous variable?


```{r}
ggplot(mpg,aes(x=displ,y=hwy))+geom_point()+facet_wrap(~cty)
```

```{r}
ggplot(mpg,aes(x=displ,y=hwy))+geom_point()+facet_grid(year~cty)
```
  
  It plots it but takes time .

2)What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))+facet_grid(drv~cyl)
```
Its shows that there are no data for such combinations hence empty cells.

3)What plots does the following code make? What does . do?


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

 "." decides which side the facet will be i.e coloumn wise or row wise ; for example (cyl~.) means facetting will be done row wise i.e cyl will be row and (.~cyl) means facetting will be done coloumn wise i.e cyl will be coloumn.


4)Take the first faceted plot in this section:

```{r}

ggplot(data = mpg) + 
geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

Ans : Advantages:
Its is better to concentrate on a particular facet alone, while with colour aesthetics is good but with large data set it will become difficult to read a particular relation .  
    Disadvantages:
 Reading global relationship is difficult with facets.
 
5)Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

```{r}
?facet_wrap
```
 
 nrow - Specifies the number of rows. 
 ncol - Specifies the number of coloumns.
 
 other layout control options:
 
 Scales: with argument 
 ="free_x"-- the x axis is free to take any scale for different facets.
 ="free_y"-- the y axis is free to take any scale for different facets.
 =fixed -- where the scales are fixed for all facets (by default)
 =free-- both x and y axes are free to take up any scale for different facets
 
 
strip.position: with arguments
 By default the labels are displayed on the top of the plot
 
"top", "bottom" "left" "right"
 
6)When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

Most screens are wider than they are tall. so it would be easier to read.

##SOME  TIPS :
##Difference between facet_wrap vs facet_grid
 https://stackoverflow.com/questions/20457905/whats-the-difference-between-facet-wrap-and-facet-grid-in-ggplot2
