---
title: ' TUTORIAL EXERCISES @ http://r4ds.had.co.nz/data-visualisation.html'
author: "Manojit Sarkar"
date: "7 July 2018"
output:
  html_document: default
  pdf_document: default
---

##Pre exercise library calls
```{r}

library(tidyverse)
```


##3.2.4 Exercises

1)Run ggplot(data = mpg). What do you see?

```{r}
ggplot(mpg)
```

Ans: No graph plot shown.

2)How many rows are in mpg? How many columns?

```{r dimension mpg}
dim(mpg)
```
Ans : 234 rows and 11 coloumns.

3)What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg
```

Ans :f = front-wheel drive, r = rear wheel drive, 4 = 4wd (what drive the car is )

4)Make a scatterplot of hwy vs cyl.

1st code :
```{r}
ggplot(mpg,aes(x=hwy,y=cyl))+geom_point()
```

2nd Code:
```{r}
ggplot(mpg)+geom_point(mapping = aes(x=hwy,y=cyl))
```



5)What happens if you make a scatterplot of class vs drv? Why is the plot not useful?


```{r}
ggplot(mpg,aes(x=class,y=drv))+geom_point()
```

Ans: The plot is not useful as the variables used are categorical variables.


## 3.3.1 Exercises
1)What’s gone wrong with this code? Why are the points not blue?


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```


Ans:Since colour = "blue" was included within the mapping argument, it was treated as an aesthetic (a mapping between a variable and a value). The expression, color="blue", treats "blue" as a variable with only one value: "blue". If this is confusing, consider how colour = 1:234 or colour = 1 would be interpreted by aes(). 
@ ans from  https://github.com/jrnold/r4ds-exercise-solutions/blob/master/visualize.Rmd

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```


2)Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

Ans there are many ways to know this 

1st way 

```{r}
sapply(mpg,class)
```

2nd way 

```{r}
? mpg 
## and then check the documentation
```

<UL>
<h1> Categorical Variables</h1>
<li>manfacturer</li>
<li>model</li>
<li>trans</li>
<li>drv</li>
<li>fl</li>
<li>class</li>

</UL>

<UL>
<h1> Continous Variables</h1>
<li>displ</li>
<li>hwy</li>
<li>year<li>
<li>cty</li>
</UL>

3)Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

## COLOR
```{r}
ggplot(mpg,aes(x=cty,y=displ,color=hwy))+geom_point()
```

## SIZE
```{r}
ggplot(mpg,aes(x=cty,y=displ,size=hwy))+geom_point()
```

## SHAPE
```{r}
#ggplot(mpg,aes(x=cty,y=displ,shape=year))+geom_point()
```

 Continous variable cannot be mapped to shape
 
## Color with categorical variable
 
```{r}
ggplot(mpg,aes(x=cty,y=displ,color=class))+geom_point()
```
Ans :For continous variable there was same colour with different intensity , while for categorical variables there are different colours for each type.


## Size with categorical variable

```{r}
ggplot(mpg,aes(x=cty,y=displ,size=fl))+geom_point()
```

A warning with size is not advised for discrete variables.


```{r}
ggplot(mpg,aes(x=cty,y=displ,shape=fl))+geom_point()
```
 It display the grap as against continous variable. But shape palette can display a maximum of 6 shapes.

3)What happens if you map the same variable to multiple aesthetics?

multi aesthetics on continous variable

```{r}
ggplot(mpg,aes(x=cty,y=displ,color=hwy,size=hwy))+geom_point()
```

yes it can be represented with color and size for continous variables.

multi aesthetics on caterogical variables.

```{r}
ggplot(mpg,aes(x=cty,y=displ,color=class,size=class))+geom_point()
```
Ans With same categorial variables can be plotted but with warnings.




4)What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}
?geom_point

ggplot(mpg,aes(x=cty,y=displ,shape=fl))+geom_point(stroke=4)
```

stroke alters the sizes of shapes


5)What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?

```{r}
ggplot(mpg,aes(x=displ,y=hwy,color=displ<5))+geom_point()
```
   
   It changes the values into boolean i.e true and false ; true / false depicting different colours each and is ploted to the graph.
   
   
   
##3.5.1 Exercises
   
1)What happens if you facet on a continuous variable?


```{r}
ggplot(mpg,aes(x=displ,y=hwy))+geom_point()+facet_wrap(~cty)
```

```{r}
ggplot(mpg,aes(x=displ,y=hwy))+geom_point()+facet_grid(year~cty)
```
  
  It plots it but takes time .

2)What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))+facet_grid(drv~cyl)
```
Its shows that there are no data for such combinations hence empty cells.

3)What plots does the following code make? What does . do?


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

 "." decides which side the facet will be i.e coloumn wise or row wise ; for example (cyl~.) means facetting will be done row wise i.e cyl will be row and (.~cyl) means facetting will be done coloumn wise i.e cyl will be coloumn.


4)Take the first faceted plot in this section:

```{r}

ggplot(data = mpg) + 
geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

Ans : Advantages:
Its is better to concentrate on a particular facet alone, while with colour aesthetics is good but with large data set it will become difficult to read a particular relation .  
    Disadvantages:
 Reading global relationship is difficult with facets.
 
5)Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

```{r}
?facet_wrap
```
 
 nrow - Specifies the number of rows. 
 ncol - Specifies the number of coloumns.
 
 other layout control options:
 
 Scales: with argument 
 ="free_x"-- the x axis is free to take any scale for different facets.
 ="free_y"-- the y axis is free to take any scale for different facets.
 =fixed -- where the scales are fixed for all facets (by default)
 =free-- both x and y axes are free to take up any scale for different facets
 
 
strip.position: with arguments
 By default the labels are displayed on the top of the plot
 
"top", "bottom" "left" "right"
 
6)When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

Most screens are wider than they are tall. so it would be easier to read.

##SOME  TIPS :
##Difference between facet_wrap vs facet_grid
 https://stackoverflow.com/questions/20457905/whats-the-difference-between-facet-wrap-and-facet-grid-in-ggplot2
 
 
##3.6.1 Exercises

What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

<B>Line Chart</B>
```{r}
ggplot(mpg,aes(x=displ,y=cty))+geom_line()
```
Ans geom_line()

<B>boxplot</B>

```{r}
ggplot(mpg,aes(x=class,y=hwy))+geom_boxplot()
```
Ans geom_boxplot()

<B>Histogram</B>

```{r}
ggplot(mpg,aes(x=displ))+geom_histogram(bins=20)
```
Ans geom_histogram()

<B>Area Chart</B>

```{r}

ggplot(mpg,aes(x=displ,y=cty))+geom_area(color="blue",fill="blue",size=1)
```

Ans; area chart

P.S : My knowledege about area chart is o far pretty limited.

Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
  
Ans : It shows a relation between engine displacement and highway milage with different colours for different drives ie front rear and 4 wd , also lines for each drives  but without the translusent interval level because of se=false.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

```

3)What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

Ans Show.legend =false hides the legends of a graph , By default show.legend is set to =TRUE.

4)What does the se argument to geom_smooth() do?

Ans se argument gives a transluent confidence interval to the geom_smooth line.

5)Will these two graphs look different? Why/why not?


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

```

```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

Ans : Exactly the same.The difference is that when the aes are set in the original ggplot, they are inherited by any other geom's that build on top of it. If you specify the aes only in a geom, it will only be used in that geom. If you use any specific aes in geom, they override the settings in ggplot.

6)Recreate the R code necessary to generate the following graphs.

##1st Graph

```{r}
ggplot(mpg,aes(x=displ,y=hwy))+geom_smooth(se=F)+geom_point()
```

##2nd Graph

```{r}
ggplot(mpg,aes(x=displ,y=hwy))+geom_smooth(se=F,aes(group=drv))+geom_point()
```

##3rd Graph

```{r}
ggplot(mpg,aes(x=displ,y=hwy,colour=drv))+geom_smooth(se=F)+geom_point()
```

## 4th Graph

```{r}
ggplot(mpg,aes(x=displ,y=hwy))+geom_smooth(se=F,)+geom_point(aes(color=drv))
```

##5 th Graph

```{r}
	
ggplot(mpg,aes(x=displ,y=hwy))+geom_smooth(se=F,aes(group=drv))+geom_point(aes(colour=drv))
```

##6th Graph

```{r}
ggplot(mpg,aes(x=displ,y=hwy))+geom_point(size=4,color="white")+geom_point(aes(color=drv))
```



##3.7.1 Exercises

1)What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

Previous Plot 
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

Ans : geop_pointrange()

```{r}
ggplot(diamonds)+geom_pointrange(aes(cut,depth),stat="summary",fun.ymax=max,fun.ymin=min,fun.y=median)
```

2)What does geom_col() do? How is it different to geom_bar()?

Ans : geom_bar uses stat_count by default: it counts the number of cases at each x position. 
geom_col uses stat_identity: it leaves the data as it is.


3)Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?


Ans  


4) What variables does stat_smooth() compute? What parameters control its behaviour?

 Ans stat_smooth() compute the y i.e predicted variable . Parameteres that control its behaviour are ymin ymax and se.


5) In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?


```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
```
```{r} 
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```


group="whatever" ( group =1 )is a "dummy" grouping to override the default behavior, which (here) is to group by "cut"" and in general is to group by the x variable. The default for geom_bar is to group by the x variable in order to separately count the number of rows in each level of the x variable. For example, here, the default would be for geom_bar to return the number of rows with cut equal to "Fair", "Good", etc.

However, if we want proportions, then we need to consider all levels of cut together.

The data are first grouped by cut, so each level of cut is considered separately. The proportion of Fair in Fair is 100%, as is the proportion of Good in Good, etc. 

group=1 (or group="anything", etc.) prevents this, so that the proportions of each level of cut will be relative to all levels of cut.


```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..,group=1))
```
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..,group=color))
```


##3.8.1 Exercises


1.What is the problem with this plot? How could you improve it?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()

```

Ans : Some of the points are overlaping each other and hene hiding the points , the solution to this is adding position: jitter and making the points a bit transperent 


```{r}
ggplot(data = mpg,aes(x = cty, y = hwy)) + geom_point(position="jitter",alpha=.5)

```

2) What parameters to geom_jitter() control the amount of jittering?

width	
Amount of vertical and horizontal jitter. The jitter is added in both positive and negative directions, so the total spread is twice the value specified.

height	
Amount of vertical and horizontal jitter. The jitter is added in both positive and negative directions, so the total spread is twice the value specified.



```{r}
ggplot(mpg, aes(x = cty, y = hwy)) + geom_point() +geom_jitter(width = 4, height = 4)
```

3)Compare and contrast geom_jitter() with geom_count().

Seems like geom_count() increases the size of the points when there are more overlapping points. Kind of like estimating the density of points in that location,also providing a legend. 
geom_jitter() just makes all the points visible by giving them a random noise and keep the points the same size.

geom_count

```{r}
ggplot(mpg, aes(x = cty, y = hwy)) + geom_count() 
```

geom_jitter

```{r}
ggplot(mpg, aes(x = cty, y = hwy)) + geom_jitter()
```
What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

```{r}
ggplot(mpg,aes(class,hwy))+geom_boxplot()
```
 Ans : the default position is "dodge"
 
 
##3.9.1 Exercises

1)Turn a stacked bar chart into a pie chart using coord_polar().

Stacked bar chart
```{r}
ggplot(diamonds,aes(x=cut,fill=clarity))+geom_bar()
```

Converting it into a Coxcomb chart using coord_polar()
```{r}
ggplot(diamonds,aes(x=cut,fill=clarity))+geom_bar(width=1)+coord_polar()
```
 
2)What does labs() do? Read the documentation.


```{r}
?labs
```

Ans It stands for label.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(position = "jitter")+labs(title="New Title",x= " New X Title",y="New Y Title")
```

3)What’s the difference between coord_quickmap() and coord_map()?

coord_map projects a portion of the earth, which is approximately spherical, onto a flat 2D plane using any projection defined by the mapproj package. Map projections do not, in general, preserve straight lines, so this requires considerable computation. coord_quickmap is a quick approximation that does preserve straight lines. It works best for smaller areas closer to the equator.


4)What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

Ans: Its shows a positive linear relationship between cty and hwy i.e as the value of cty increases  hwy also increases.
The relationships is approximately linear, though overall cars have slightly better highway mileage than city mileage. 

coord_fixed()
A fixed scale coordinate system forces a specified ratio between the physical representation of data units on the axes. The ratio represents the number of units on the y-axis equivalent to one unit on the x-axis. The default, ratio = 1, ensures that one unit on the x-axis is the same length as one unit on the y-axis. Ratios higher than one make units on the y axis longer than units on the x-axis, and vice versa.

geom_abline()
draws a line that, by default, has an intercept of 0 and slope of 1.
