---
title: "Chapter4 Workflow Basics"
author: "Manojit Sarkar"
date: "10 July 2018"
output: html_document
---

1)Why does this code not work?
```{r}
my_variable <- 10
#my_varlable
```
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found

Ans : There is a typing error it should be my_variable instead of my_varlable
```{r}
my_variable
```

2)Tweak each of the following R commands so that they run correctly:


```{r}
library(tidyverse)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
#fliter(mpg, cyl = 8)
```


```{r}
#filter(diamond, carat > 3)
```

Ans : 

```{r}
library(tidyverse)
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))
```

flitre instead of filtre

```{r}
filter(mpg,cyl==8)
```

data =diamond instead of diamonds

```{r}
filter(diamonds, carat > 3)
```

3)Press Alt + Shift + K. What happens? How can you get to the same place using the menus?

 It opens up a Key Board Shortcut Quick Reference window.
 
We can go to the same window through menus ---
Tools > Keyboard Shortcut Help